### 2024-04-17 paper/paper.qmd

---
title: "An Analysis of the Demographics of Toronto's Homeless Population and the Relationship Between the Number Deaths of Homeless Individuals Based on the Month"
author: 
  - Irene Huynh
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "Toronto’s shelter system and homeless population has experienced a significant increase over recent years. Using Toronto’s shelter system data and deaths of people experiencing homelessness data from the City of Toronto's Open Data Portal, this paper analyzes when deaths most frequently occur in Toronto’s homeless population and the trends of different demographics that utilized shelters from 2018 to 2023. This paper studies the month of death of unhoused individuals, the age group that visits shelters the most, the trend of the average number of individuals who are actively homeless using shelters, the average number of people who return to shelters after acquiring housing, and the gap separating the average number of male and female shelter users from 2018 to 2023. This paper finds that deaths of unhoused individuals occur most frequently in January and July. Additionally, males and individuals aged 25 to 44 account for a significantly larger portion of shelter users. This study may be employed to inform policy makers of the demographic most at risk of homelessness and of where to target solutions for tackling the underlying issue of homelessness in Toronto."
format: 
  pdf:
    fig-pos: 'H'
header-includes:
  - \usepackage{float} 
  - \floatplacement{table}{H}
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)

cleaned_toronto_shelters_data <- read_csv("../data/analysis_data/cleaned_toronto_shelters_data.csv")
```

# Introduction

Toronto’s shelter system provides a temporary home to those who are in need of one. The system is essential for Toronto’s homeless population and refugees as the shelters uphold standards that support meals and laundry accommodations, mental health and harm reduction services, and a host of counselors to direct and help individuals acquire permanent housing. Although the shelters provide a multitude of services and aid to secure permanent housing for the unhoused, there are still many who return to shelters after securing housing [@torontoshelters].

The affordable housing crisis is directly contributing to Toronto’s shelter availability crisis. As Toronto’s housing market has skyrocketed in the past few years, it is no surprise that Toronto’s shelters are receiving significantly more visitors. The City of Toronto states that as of 2023, “One in three of Toronto’s homeowner and renter households (373,965 households or 32 per cent) experienced affordability issues,” [@cityToronto]. According to CBC News, an estimated average of at least 273 individuals were turned away from shelters in June 2023, which was a record high for Toronto shelters [@cbcshelters]. The amount of families seeking warm beds has been unmatched and as Toronto’s winters can be extremely frigid, shelter availability is becoming increasingly more critical. Moreover, as temperatures continue to reach record highs as a consequence of climate change, Toronto’s summers become increasingly more hazardous for unhoused individuals.

Analyzing and studying Toronto’s shelter system data can reveal those within society that are struggling the most with securing housing. This paper analyzes the demographics that utilize Toronto’s shelter system and the month of deaths of unhoused individuals. Analyzing the deaths of individuals experiencing homelessness is of particular interest to produce a predictive model that determines whether specific months impact the likelihood of death in Toronto’s homeless population. In addition, this paper is also concerned with the actively homeless population and different demographics, such as refugees and single adults, that stay in shelters. The main estimand of this study is the effect of months on deaths in individuals experiencing homelessness in Toronto. This paper finds that the largest demographic of shelter users are males and individuals aged 25 to 44, and the death of people experiencing homelessness occurs most frequently in January and July.

The remainder of this paper is structured as follows: the Data section details the collection
and processing of Toronto shelter system and deaths of unhoused individuals data. Additionally, the section includes discussion on the measurement of the original dataset and the variables examined in this study. The following Models section discusses the regression model used and the Results section displays the tables and figures that model the data. The Discussion section draws conclusions from those tables and figures and details the importance of my findings and weaknesses of the paper.

# Data

## Sources

The data utilized in this paper was retrieved from the City of Toronto's Open Data Portal [@opendatatoronto]. The statistical programming language R was used to retrieve, clean, and process the data [@citeR]. In particular, the following R packages were used: `opendatatoronto` [@opendatatoronto], `tidyverse` [@tidyverse], and `dplyr` [@dplyr] for data acquisition; `janitor` [@janitor] and `arrow` [@arrow] for data cleaning and processing; `rstanarm` [@rstanarm], `marginaleffects` [@marginaleffects], and `knitr` [@knitr] for modeling; and `ggplot2` [@ggplot2] for creating figures.

## Toronto Shelter System Data

The shelter system data acquired from the City of Toronto's Open Data Portal provided monthly statistics of Toronto's shelters from January 2018 to December 2023. For each month, the data set included information for the entire population, chronic visitors, refugees, families, youth, single adults, and non-refugees. The information of interest for each of those demographics are the number of individuals that returned to shelters after previously acquiring housing, those that are new to the shelter system, those that are actively homeless, those that identify as male and female, and the number of individuals in each age group in that demographic. The age groups consist of visitors under 16, 16 to 24, 25 to 44, 45 to 64, and 65 years of age or older. 
## Deaths of People Experiencing Homelessness Data

The death count for individuals experiencing homelessness data acquired from the City of Toronto’s Open Data Portal provided monthly statistics of deaths from January 2017 to June 2023.

## Measurement

The City of Toronto’s Open Data Portal is Toronto's official source for city data relating to each of its divisions and agencies. Thus, the quality of the data used in this study is reliable; however, the Portal does state that there were limitations to the data sets. For the Toronto shelter system data, a limitation or possible avenue for measurement error for the data is that the data excludes individuals who do not sleep in shelters or who use other resources for shelter, such as public infrastructure and buildings open all hours  [@torontosheltersystemflow]. Additionally, for the deaths of people experiencing homelessness data, the data only includes reported deaths [@homelessdeathsdata]. Therefore, the number of deaths could be much higher as unreported deaths were excluded from the dataset.

# Model

The goal of my modelling strategy is to utilize the Poisson distribution to build a Poisson regression model as the data is concerned with count data and $\lambda$ distributes probabilities over non-negative integers.

## Model set-up

Define $y_i$ as the number of deaths and the explanatory variable as the month. I am interested in how the number of deaths differs by month.

\begin{align} 
y_i|\lambda_i &\sim \mbox{Poisson}(\lambda_i) \\
\mbox{log}(\lambda_i) &= \beta_0 + \beta_1 \times \mbox{Month}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\end{align}

I run the model in R [@citeR] using the `rstanarm` package of @rstanarm. I use the default priors from `rstanarm`.

# Results

## Relationship Between Deaths of Homeless Individuals and Months
```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/deaths_months_prediction.rds"))
```

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Examining the number of deaths of homeless individuals occuring in each month"
#| tbl-pos: H
#| echo: false
#| warning: false
#| eval: true

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

## Relationship Between Deaths of Homeless Individuals and Months

@tbl-modelresultsslopes helps with interpreting the model results in @tbl-modelresults. @tbl-modelresultsslopes suggests that in our dataset, compared to April, August tends to have around three additional deaths, December tends to have around two additional deaths, February tends to have one additional death, and so on. Most notably, compared to April, January tends to have five additional deaths–the most additional deaths out of all the months; July tends to have around four additional deaths–the second most additional deaths out of all the months; and March tends to have one less death–the least amount of additional deaths out of all the months.

```{r}
#| label: tbl-modelresultsslopes
#| tbl-cap: "The estimated difference in the number of deaths of homeless indivuals in each month"
#| tbl-pos: H
#| echo: false
#| warning: false
#| eval: true

library(knitr)
library(marginaleffects)

slopes(first_model) |>
  select(contrast, estimate) |>
  unique() |> 
  kable(
    col.names = c("Compare month", "Estimate"),
    digits = 2, booktabs = TRUE, linesep = ""
  )
```

```{r}
#| label: fig-modelresultsfig
#| fig-cap: "The predicted number deaths of homeless individuals based on the month"
#| echo: false
#| warning: false
#| eval: true

#library(marginaleffects)

#plot_predictions(first_model, condition = "month_of_death") +
#  labs(x = "Month",
#       y = "Number of Deaths") +
#  theme_classic()
```

## Average Number of Indivduals for each Age Group

@fig-agegroupavg demonstrates that individuals age 25 to 44 enter shelters the most on average, followed by individuals age 45 to 64, then those under 16 and those age 16 to 24. Additionally, individuals 65 and over enter shelters the least on average. This result is expected as the cost of living in Toronto has risen significantly in recent years and individuals age 25 to 44 must have a large percentage of their income allotted to rent [@cbcliving]. This quota may be difficult to reach every month as the price of other necessities have also risen, such as grocery bills. Individuals over 65 may not choose to stay in shelters as they have other options, including retirement homes and elder care centres.

```{r}
#| label: fig-agegroupavg
#| fig-cap: "The average of all shelter visitors in each age group over 6 years from 2018 to 2023"
#| echo: false
#| warning: false
#| message: false

age_group_data <-  cleaned_toronto_shelters_data|>
    filter(population_group == "All Population") |>
  summarise(
    # Divide by (6*12) since it is 6 years (2018-2023) with 12 months each
    "Under 16" = sum(age_under16, na.rm = TRUE)/(6*12), 
    "16-24" = sum(age_16_24, na.rm = TRUE)/(6*12),  
    "25-44" = sum(age_25_44, na.rm = TRUE)/(6*12),   
    "45-64" = sum(age_45_64, na.rm = TRUE)/(6*12), 
    "65+" = sum(age_65over, na.rm = TRUE)/(6*12)
  ) |>
  pivot_longer(cols = everything(), names_to = "age_group", values_to = "average")

# Create the bar graph
ggplot(age_group_data, aes(x = age_group, y = average, fill = age_group)) +
  geom_bar(stat = "identity") +
  labs(x = "Age Group",
       y = "Average")
```

## Average Number of Actively Homeless Indivduals

@fig-activelyhomeless demonstrates that the average number of actively homeless individuals increased from 2018 to 2019, significantly decreased from 2019 to 2020, remained almost unchanged from 2020 to 2021, experienced a sharp increase from 2021 to 2022, and increased more from 2022 to 2023. Additionally, the average in 2023 reached a high of almost 10500 actively homeless individuals. The housing crisis and spike in the cost of living since the COVID-19 pandemic is a notable contributing factor to the increase in homelessness and shelter usage after 2021. The significant decrease in the average number of actively homeless individuals using shelters from 2019 to 2021 can also be attributed to the pandemic: many public areas were restricted or fully closed for physical distancing purposes.

```{r}
#| label: fig-activelyhomeless
#| fig-cap: "The average number of actively homeless people from 2018 to 2023"
#| echo: false
#| warning: false
#| message: false

avgs_actively_homeless <- rep(NA, times=6) # 6 years (2018 to 2023)
year = 18
for (i in 1:6){
  each_year_actively_homeless <- cleaned_toronto_shelters_data |>
    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
    select(actively_homeless)

  avgs_actively_homeless[i] <- sum(each_year_actively_homeless)/12 # 12 months in a year

  year = year + 1
}

avgs_actively_homeless_data <-
  tibble(
    date = 2018:2023,
    avg = avgs_actively_homeless
  )

ggplot(avgs_actively_homeless_data, aes(x = date, y = avg, fill = date)) +
  geom_line(stat = "identity") +
  labs(x = "Year",
       y = "Average")
```

## Average Number of Shelter Returners

@fig-returnedfromhousingavg illustrates that the average number of individuals who return to shelters each year increased each year from 2018 to 2021, then the average began to decrease from 2021 to 2022 and significantly declined from 2022 to 2023. In addition, less than 60 individuals returned to a shelter in 2023. This outcome is surprising as it was expected to find that after acquiring housing, many would return to shelters due to the rise of rental prices and the overall housing and cost of living crisis in recent years. Additionally, @fig-activelyhomeless demonstrates the significant increase in homeless individuals. A reason that can explain why many homeless people who acquire housing and become homeless again but do not return to shelters is that more individuals are experiencing homelessness; thus, shelter capacity is quickly filling up and many returning visitors are being turned away [@cbcshelters].

```{r}
#| label: fig-returnedfromhousingavg
#| fig-cap: "The average of number of visitors who returned to shelters from 2018 to 2023"
#| echo: false
#| warning: false
#| message: false

avgs_returned <- rep(NA, times=6) # 6 years (2018 to 2023)
year = 18
for (i in 1:6){
  each_returned_from_housing <- cleaned_toronto_shelters_data |>
    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
    select(returned_from_housing)

  avgs_returned[i] <- sum(each_returned_from_housing)/12 # 12 months in a year

  year = year + 1
}

avgs_returned_data <-
  tibble(
    date = 2018:2023,
    avg = avgs_returned
  )

ggplot(avgs_returned_data, aes(x = date, y = avg, fill = date)) +
  geom_line(stat = "identity") +
  labs(x = "Year",
       y = "Average")
```

## Average Number of Males and Females

There is an average of over 5000 male individuals in shelters each year (blue); whereas for female individuals (red), there is an average of less than 4000. @fig-malefemale illustrates how every year, the average number of male shelter users is significantly higher than that of female users. Additionally, from 2020 to 2021, the average number of male shelter users increased while the average number of female shelter users decreased. In regards to other annual intervals, the average for males and females both increase from 2018 to 2019, decrease from 2019 to 2020, and increase from 2021 to 2023.

```{r}
#| label: fig-malefemale
#| fig-cap: "The average of male and female shelter visitors from 2018 to 2023"
#| echo: false
#| warning: false
#| message: false

avgs_male <- rep(NA, times=6) # 6 years (2018 to 2023)
avgs_female <- rep(NA, times=6)
year = 18
for (i in 1:6){
  each_year_male <- cleaned_toronto_shelters_data |>
    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
    select(gender_male)

  each_year_female <- cleaned_toronto_shelters_data |>
    filter(population_group == "All Population", substring(date, 5, 6) == as.character(year)) |>
    select(gender_female)

  avgs_male[i] <- sum(each_year_male)/12 # 12 months in a year
  avgs_female[i] <- sum(each_year_female)/12

  year = year + 1
}

avgs_gender_data <-
  tibble(
    date = 2018:2023,
    avg_males = avgs_male,
    avg_females = avgs_female
  )

ggplot(avgs_gender_data, aes(x=date)) +
  geom_line(aes(y = avg_males), color="blue") +
  geom_line(aes(y = avg_females), color="red")+
  labs(x = "Year",
       y = "Average")
```

# Discussion

## Findings

This study utilized a Poisson regression model to predict the number of deaths of homeless individuals based on the month. The analysis conducted in this study found that the most deaths occurred in January, the second most deaths occurred in July, and the least deaths occurred in March. These findings are expected considering that January can consist of Toronto’s most frigid, cold weather, causing homeless individuals who cannot secure shelter to suffer in the extremely cold temperatures. July consists of the other extreme weather case: severe heat waves, which can be fatal. This study also found that the most prevalent age group that uses shelters is the 25 to 44 age range and the number of males that use shelters significantly surpass the number of females. Additionally, the number of actively homeless people in Toronto is trending upwards as of 2021.

## Importance

Housing is becoming increasingly difficult to acquire in Toronto, leaving thousands unhoused as temperatures reach their extremes in January and July. Many factors can be to blame, including the cost of living crisis, the housing crisis, and the COVID-19 pandemic. According to a new report, “The cost of living in Toronto went up nearly 20 percent between 2017 and 2022–but that’s less than half the increase seen in the city’s housing prices over the last five years,” [@cp24]. As the cost of necessities soar, meeting rent payments becomes exceedingly difficult to accomplish, causing more and more people to find themselves unhoused. Homeless individuals are in vulnerable positions, but are often stigmatized as they face discrimination as a direct result of their housing status. Professionals and businesses are less willing to offer their services, leading to higher rates of poor physical and mental health. According to @stigma, there is evidence that “homeless individuals are not perceived as fully human,” making it tremendously distressing when searching for employment, healthcare, and sympathy. As much of society vilifies the unhoused, it is important to identify ways to provide relief for them. This paper reveals that deaths in Toronto’s homeless population occur most often in the coldest and warmest months of the year. With this knowledge, the City of Toronto can establish more temporary shelters during January and July and if that is not feasible, funding can be directed to establishing warming stations that can quickly be erected in January and to establishing cooling stations in July.

## Weaknesses and Future Proceedings

Although this paper used data from a reliable source, there were some limitations to the collection of the data which became potential sources for measurement error. Such limitations included the exclusion of unreported deaths of homeless individuals and shelters that are not part of the City of Toronto’s database. Moreover, the data used did not include the cause of death of individuals, leading to the speculation that the weather is solely responsible for the deaths; however, there are many other possibilities, such as underlying health conditions or drug overdoses. In future studies of homelessness and shelter usage in Toronto, an analysis of the cause of death of homeless individuals would provide more insight into where to target relief efforts.

\newpage

# References